\input{../preamble.tex}
\usepackage{indentfirst}

% Title page stuff
\title{\textit{D}-Modules Over Smooth Affine Varieties}
\date{}
%\author{Author Name}
%\use package[show frame]{geometry}

\newcounter{entry-counter}
\newcommand{\entry}[1]
{
	\addtocounter{entry-counter}{1}
    \tchap{Entry \arabic{entry-counter}}
	\addcontentsline{toc}{section}{Entry \arabic{entry-counter}: #1}
	\addtocounter{section}{1}
	\vspace{-1.5em}
    \begin{center}
		\small \emph{Written: #1}
    \end{center}
}

\usepackage{titling}
\renewcommand\maketitlehooka{\null\mbox{}\vfill}
\renewcommand\maketitlehookd{\vfill\null}

% start document
\begin{document}
\pagestyle{empty}
	\LARGE
\begin{center}
	Sporadic Notes Taken While at LPS Probabilistic Computing \\
	\Large
	Isaac Martin \\
    Last compiled \today
\end{center}
\normalsize
\vspace{-2mm}
\hru
\tableofcontents

\entry{2022-Aug-17}

\textbf{Goals}: Know your people. Get to know Tim, Liesbeth, the assistant graduate professor Maria Gualdani (all things academic), your feloow graduate students (especially older grad students), your instructors and professors, and of course your advisor.

Know UT Resources (MyUT, math website, Graduate School website, Texas One Stop).

Know when to ask: ask/share before things become an actual issue or a burden or a problem.

\bigskip

\textbf{Navigating UT Resources:}
\begin{itemize}
	\item MyUT: my.utexas.edu. Contains registration information sheet, registration times, how to register, etc. Finances: Direct Deposit, earning statements. Campus resources, IT to campus saftey and public transportation. Look at \textbf{Tuition Bill} (deadline to confirm registration) and \textbf{What I Owe Page} to see what refunds you may get. Always look at the Tuition Bill.
	\item Want to do advising with the graduate advisor and grad coordinator during the first year at some point, will be advised prior to every semester in the first year.
	\item Math website contains student handbook, guidance on processes, topics courses, etc.
\end{itemize}

\bigskip

What is the Graduate Coordinator?
\begin{itemize}
	\item Assists in anything related to administration.
	\item Provides info on policies and procedures pertaining to the University, Graduate School and math department
	\item Supports students in collaboration with the graduate advisor
	\item Seems absurdly friendly and wants to help everybody with everything so that's pretty sick
	\item Hybrid work schedule (2 days in-office, Tue/Wed -- 3 days remote)
	\item Office: PMA 8.144 -- open door policy i.e. if they're in office then the door is open.
\end{itemize}

Code for important room whose number I forgot is 13579\#. It has paper and a printer and stuff.

\bigskip

\textbf{Registering} always follows the same pattern.
\begin{itemize}
	\item Course schedule is published (mid semester) and you identify courses to take.
	\item You get advised
	\item You solve your registration bars (advising bar is one that Liesbeth lifts)
	\item Register for classes during the 1st or 2nd access period (1st period is preferred if possible, June/July for Fall October/...November(?) for Spring). Vital to register in first period if you're on fellowship funding.
	\item Visit Texas One Stop or MyUT to see payment options
	\item Wait until you see that all deductions have been applied to your bill
\item Pay your tuition bill OR confirm by the deadline (If you have a ZERO DOLLAR bill, you \red{MUST still confirm registration!}) 
\end{itemize}

\bigskip

\textbf{What is your registration information sheet (RIS)?} It shows
\begin{itemize}
	\item Semester to register
	\item Personal info
	\item Student health Insurance plan
	\item Your \red{access time to register} (vital)
	\item Access time to add/drop courses
	\item Registration bars
\end{itemize}

\textbf{Student Employment:} (TA/AI/GRA)
\begin{itemize}
	\item \textbf{Academic Graduate Students Employees (AGS)} appointment dates are Fall: 9/1 - 1/15, Spring: 1/16-5/31, Summer: 6/1 - 8/31
	\item Workday: UT's HR system. Onboarding and other requirements need to be done promoptly.
\end{itemize}

\bigskip

\textbf{Graduate Student Fellows}
\begin{itemize}
	\item Also in Workday as nonemployee (graduate student fellow)
	\item Certain fellowships have (UT) work restrictions -- DON'T SNEAK
	\item Payment schedule is different than student employment. First payment for fall is Sept 1. IF YOU RECEIVE A DOUBLE PAYMENT, it's because you're not going to be paid for a pay period. Save it don't spend it.
\end{itemize}

\bigskip

\textbf{AcademicBlue SHIP}: This is for insurance or something. Fellowship students use the ''UT Austin Voluntary and UT Austin International and ESL`` button to check policy and self-enroll in the health insurance. International students and TA/AI/GRA students do not self-enroll, they use this page exclusively to access the policy.

\bigskip

You're going to get a lot of emails from UT, but always read emails from Maria, Liesbeth, Tim and those pertaining to health insurance and payment.

\bigskip

\textbf{Deadlines to remember:}
\begin{itemize}
	\item Tuition and fee payment \red{due by 5:00 pm, August 25}
	\item \red{Everybody needs to log in to their tuition bill and take action by the deadline.}
	\item The twelfth day of class, \red{September 10} is the official enrollment count day, is the last day to add a class, and is the last day to drop a class for a possible refund.
\end{itemize}

\begin{itemize}
	\item UT ID card - FAC building (student ID center) public transportation is free with that card
	\item Keep your contact info up to date in MyUT - directory
	\item Update/acknowledge your emergyency contact information every semester, it's a bar on registration I believe
	\item Set up Direct Deposit
	\item Parking on UT Campus -- see Parking Profile to see what options you have. You have access to different parking office when holding student employment
	\item Sign Milestone agreement
	\item Key slips -- Locks and Keys
	\item Contacting Grad School -- cc your graduate coordinator
	\item Read our emails! -- weekly gRAD MAIL
	\item Use of listserve
	\item UTBox Student folders
\end{itemize}


\printbibliography
\end{document}
