\input{../../preamble.tex}

\begin{document}
\pagestyle{empty}
	\LARGE
\begin{center}
	Problems from Hartshorne Chapter 2.2 \\
	\Large
	Isaac Martin \\
    Last compiled \today
\end{center}
\normalsize
\vspace{-2mm}
\hru
\begin{homework}[e]
	\prob[\textsc{Exercise 2.7.}] Let $X$ be a scheme. For any $x \in X$, let $\cO_x$ be the local ring at $x$, and $\frakm_x$ its maximal ideal. We define the \emph{residue field} of $x$ on $X$ to be the field $k(x) = \cO_x/\frakm_x$. Now let $K$ be any field. Show that to give a morphism of $\Spec K$ to $X$ it is equivalent to give a point $x \in X$ and an inclusion map $k(x) \to K$.
	\begin{prf}
		Suppose first that we have a map $f:\Spec K \to X$. Topologically, this is determined solely by choosing an image $x\in f(P)$ for the sole point $P\in \Spec K$. Sheaf theoretically, this consists of a map $f^\sharp:\cO_X \to f_*\cO_K$ (by $\cO_K$ we mean $\cO_{\Spec K}$). This induces a local ring map on the stalk at $P$: $f^\sharp_P:\cO_{X,x}\to (f_*\cO_{K})_P = K$, meaning that the maximal ideal $\frakm_x$ in $\cO_{X,x}$ is sent to the maximal ideal $(0) \subseteq K$, meaning that $\frakm_x = \ker f^\sharp_P$. This in turn implies that $f^{\sharp}_P$ factors through the quotient $\pi:\cO_{X,x} \mapsto k(x) = \cO_{X,x}/\frakm_x$ and hence induces a map $k(x) \to K$. This map is necessarily an inclusion since every ring homomorphism of fields is injective.

		Now suppose we have an injection $p:k(x) \hookrightarrow K$. We can then define a map $f^\sharp_x:\cO_{X,x}\to K$ by $f^{\sharp} = p\circ \pi$, where $\pi:\cO_{X,x}\to k(x)$ is the quotient map. This is precisely a map on between the stalks $\cO_{X,x}$ and $\cO_{K,P}$. If we define $f:\Spec K\to X$ by $P\mapsto x$ and $f^\sharp(U):\cO_X(U) \to f_*\cO_K(U) = K$ by $f^\sharp(U) = f^{\sharp}_x \circ \iota$ where $\iota:\cO_X(U) \to \cO_{X,x}$ is the natural localization map, then $(f,f^\sharp)$ is a map of schemes. Note that for any open set $U \subseteq X$ not containing $x$ the map $f^\sharp:\cO_X(U)\to f_*\cO_K(U)$ is necessarily the zero map, since $f_*\cO_K(U) = \cO_K(f^{-1}(U)) = \cO_K(\emptyset) = 0$.
	\end{prf}
	\prob[\textsc{Exercise 2.11.}] Let $k = \bF_p$ be the finite field with $p$ elements. Describe $k[x]$. What are the residue fields of its points? How many points are there with a given residue field?
	\begin{prf}
		The ring $k[x]$ is a PID since $k$ is a field, so the prime ideals are all principally generated by irreducible polynomials $f \in k[x]$.
	\end{prf}
\end{homework}
\end{document}
