\input{../preamble.tex}

% lecture commands
\newcounter{lecture-counter}
\newcommand{\lecture}[2]
{
	\newpage
	\addtocounter{lecture-counter}{1}
    \tchap{Lecture \arabic{lecture-counter}}
	\vspace{-1.5em}
    \begin{center}
	    \small \emph{Recorded: #1 \hspace{1.5em} Notes: #2}
    \end{center}
}

% start document
\begin{document}
\pagestyle{empty}
	\LARGE
\begin{center}
	Lecture Notes from Differential Geometry (Michaelmas 2021) \\
	\Large
	Isaac Martin \\
    Last compiled \today
\end{center}
\normalsize
\vspace{-2mm}
\hru
\tableofcontents

\lecture{2022-03-09}{2022-03-09}

\begin{cor}\label{cor:roots-of-unity-and-n-torsion}
	If $E[n] \subseteq K(K)$ then $\mu_n \subseteq K$, where $\mu_n$ is the set of $n$th roots of unity in $\ol{K}$.
\end{cor}
\begin{prf}
	If $e_n$ is nondegenerate then there exist $S,T \in E[n]$ such that $e_n(S,T)$ is a primitive $n^{th}$ root of unit, say $\zeta_n$. Then $\sigma(\zeta_n) = e_n(\sigma S,\sigma T) = e_n(S,T) = \zeta_n$ for all $\sigma \in \Gal(\olK/K)$. The first equality follows from Galois equivalence and the second since $S,T \in E(K)$. Therefore $\zeta_n \in K$.
\end{prf}
\begin{example}
	There exists no $E/\bQ$ such that $E(\bQ)_{tors} \cong (\bZ/3\bZ)^2$.
\end{example}
\begin{rmk}
	In fact, the Weil pairing is alternating, i.e. $e_n(T, T) = 1$ for all $T \in E[n]$. In particular, expanding $e_n(S+T,S+T)$ show $e_n(S,T) = e_n(T,S)^{-1}$.
\end{rmk}
\section{Galois Cohomology}
Throughout this section, $G$ is a group and $A$ is a $G$-module, i.e. and abelian group with an action of $G$ via group homomorphisms. That is, we have a map $G \to \Aut(A)$ where $\Aut(A)$ is the group of abelian group homomorphisms of $A$, and $g\cdot a = g(a)$. To say that $A$ is a $G$=module is equivalent to saying that $A$ is a $\bZ[G]$-module.

\begin{defn}\label{defn:basic-homology-defs}
	We set
	\begin{align*}
		H^0(G,A) = A^G = \{a \in A ~\mid~ \sigma(a) = a, ~ \forall \sigma \in G\}.
	\end{align*}
	We further set
	\begin{align*}
		C^1(G,A) &= \{\text{maps } G\to A\} & \text{``cochains''} \\
		Z^1(G,A) &= \{(a_\sigma)_{\sigma \in G} ~ \mid ~ a_{\sigma \tau} = \sigma(a_\tau) + a_\sigma\} & ``cocycles'' \\
		B^1(G,A) &= \{(\sigma b - b)_{\sigma \in G} ~ \mid ~ b \in A\} & \text{``coboundariers''}
	\end{align*}
	and we have inclusions $B^1(G,A) \subseteq Z^1(G,A) \subseteq C^1(G,A)$. We define $H^1(G,A) = Z^1(G,A)/B^1(G,A)$.
\end{defn}
\begin{rmk}
	If $G$ acts trivially on $A$, then $H^1(G,A) = \Hom(G,A)$.
\end{rmk}
\begin{thm}\label{thm:ses-induces-les-on-cohomology}
	A short exact sequence of $G$-modules
	\begin{align*}
		0 \to A \to{\phi} B \to{\psi} C \to 0
	\end{align*}
	gives rise to a long exact sequence of abelian groups
	\begin{align*}
		0 \to A^G \to[\phi] B^G \to[\psi] C^G \to[\delta] H^1(G,A) \to[\phi_*] H^1(G,B) \to[\psi_*] H^1(G,C) \to ...
	\end{align*}
	where we stop before $H^2(G,A)$ because we have yet to define it. The map $\delta$ arises from the snake lemma.
\end{thm}
\begin{defn}\label{defn:delta-in-les-cohomology}
	Let $c \in C^G$. Then there exists a $b \in B$ such that $\psi(b) = c$. Then
	\begin{align*}
		\psi(\sigma b - b) = \sigma(c) - c= 0
	\end{align*}
	for all $\sigma \in G$. This means $\sigma b - b = \phi(a_\sigma)$ for some $a_\sigma \in A$. One checks that $(a_\sigma)_{\sigma \in G} \in Z^1(G,A)$. We define $\delta(c) = \text{ chars of } (a_\sigma)_{\sigma \in G}$ in $H^1(G,A)$.
\end{defn}
\begin{thm}\label{thm:normal-subgroup-exact-sequence}
	Let $A$ be a $G$-module $H \subseteq G$ a normal subgroup. Then there is an inflation-restriction exact sequence
	\begin{align*}
		0 \to H^1(G/H,A^H) \to{\inf} H^1(G,A) \to{\res} H^1(H,A)
	\end{align*}
\end{thm}
\begin{prf}
	Omitted.
\end{prf}
Let $K$ be a perfect field. $\Gal(\olK/K)$ is then a topological group with basis of open subgroups. The sets $\Gal(\olK/L)$ for $[L:K] < \infty$.

If $G = \Gal(\olK/K)$ then we modify the definition of $H^1(G,A)$ by insisting
\begin{enumerate}
	\item The stabilizer of each $a \in A$ is an open subgroup of $G$.
	\item All cochains $G \to A$ are continuous where $A$ is given by the discrete topology.
\end{enumerate}
Then
\begin{align*}
	H^1(\Gal(\olK/K),A) = \varinjlim_{L, ~L/K \text{finite Galois}} H^1(\Gal(L/K), A^{\Gal(\olK/L)}).
\end{align*}
The direct limit is with respect to inflation maps (what are inflation maps?).

\begin{thm}[Hilbert's Theorem 90]\label{thm:Hilbert90}
	Let $L/K$ be a finite Galois extension. Then $H^1(\Gal(L/K),L^*) = 0$.
\end{thm}
\begin{prf}
	Let $G = \Gal(L/K)$. Let $(a_\sigma)_{\sigma \in G} \in Z^1(G,L^*)$. Distinct automorphisms are linearly independent, hence there exists some $y \in L$ such that
	\begin{align*}
		\underbrace{\sum_{\tau \in G} a_\tau^{-1} \tau(y)}_x \neq 0.
	\end{align*}
	For $\sigma \in G,$
	\begin{align*}
		\sigma(x) = \sum_{\tau \in G} \sigma(a_\tau)^{-1} \sigma \tau(y) = a_\sigma \sum_{\tau \in G} a_\sigma^{-1} \sigma \tau (y) = a_\sigma \cdot x.
	\end{align*}
	Therefore $a_\sigma = \sigma(x)/x \implies (a_\sigma)_{\sigma \in G} \in B^1(G,L^*)$. Hence $H^1(G,L^*)$.
\end{prf}

\begin{cor}\label{cor:vanishing-first-cohomology}
	$H^1(\Gal(\olK/K),\olK^*) = 0$.
\end{cor}
Application: Assume $\fchar K \not | n$. There is an exact sequence of $\Gal(\olK/K)$-modules
\begin{align*}
	0 \to \mu_n \to \olK^* \to[x\mapsto x^n] \olK^* \to 0.
\end{align*}
Have a long exact sequence
\begin{align*}
	K^* \to[x \mapsto x^n] K^* \to H^1(\Gal(\olK/K),\mu_n) \to H^1(\Gal(\olK/K),\olK^*),
\end{align*}
but $H^1(\Gal(\olK/K),\olK^*) = 0$ by Theorem (\ref{thm:Hilbert90}). Therefore $H^1(\Gal(\olK/K),\mu_n) \cong K^*/(K^*)^n$. 

If $\mu_n \subseteq K$ then $\Hom_{cts}(\Gal(\olK/K),\mu_n)\cong K^*/(K^*)^n$.

If $L/K$ is a finite Galois extension then $\Gal(\olK/K) \to{\pi} \Gal(L/K)$ and hence
\begin{align*}
	\Hom(\Gal(L,K),\mu_n) \hookrightarrow \Hom_{cts}(\Gal(\olK/K),\mu_n) \cong  K^*/(K^*)^n,
\end{align*}
where the above map is given by $\chi \mapsto \chi \circ \pi$. The image is a finite subgroup $\Delta \subseteq K^*/(K^*)^n$. 

If $\Gal(L/K)$ is abelian of exponent dividing $n$ then
\begin{align*}
	[L:K] = |\Gal(L/K)| = |\Hom(\Gal(L/K),\mu_n)| = |\Delta|.
\end{align*}
Compare to Theorem 11.2 from lectures \red{Fix numbering}.

\textbf{Notation:} We'll write $H^1(K,-) = H^1(\Gal(\olK/K),-)$ to avoid writing $\Gal$ and $\olK$ every time.

\begin{lem}\label{lem:finite-padic-ext}
	Let $[K:\bQ_p] < \infty$. Then
	\begin{align*}
		\ker(H^1(K,\mu_n) \to H^1(K^{nr},\mu_n)) \subseteq \{x \in K^*/(K^*) ~ \mid~ v(x) \equiv 0 ~ (\mod n)\}.
	\end{align*}
	remember that $K^{nr}$ is the maximal unramified extension of $K$.
\end{lem}
\begin{prf}
	By Theorem (\ref{thm:Hilbert90}), identify $H^1$
\end{prf}

\lecture{2022-03-11}{2022-03-11}
\begin{lem}\label{lem:ker-of-H1}
	Let $K:\bQ_p] < \infty$. Then
	\begin{align*}
		\ker(H^1(K,\mu_n)\to H^1(K^{nr},\mu_n)) \subseteq \{ x \in K^*/(K^*)^n ~ \mid ~ v(x) \equiv 0 ~(\mod n)\}
	\end{align*}
\end{lem}
\begin{prf}
	(Continued). The discrete valuation $v: K^*\to \bZ$ extends to $v:(K^{nr)^* \to \bZ})$. Then $v(x) = nv(y) \equiv 0 ~(\mod n)$.
\end{prf}

\textbf{EXERCISE:} (in local fields.) Show that if $p \not | n$ then $\subseteq$ is actually $=$.

Let $\phi:E\to E'$ be an isogeny of elliptic curves over $K$. Then there is a short exact sequence of $\Gal(\olK/K)$-modules
\begin{align*}
	0 \to E[\phi] \to E \to{\phi} \to E' \to 0.
\end{align*}
Long-exact sequence:
\begin{align*}
	E(K) \to{\phi} E'(K) \to{\delta} H^1(K,E[\phi]) \to H^1(K,E) \to{\phi_*} H^1(K,E').
\end{align*}
We get a short exact sequence
\begin{align*}
	0 \to \frac{E'(K)}{\phi E(K)} \to H^1(K,E[\phi]) \to H^1(K,E)[\phi*] \to 0.
\end{align*}
Now take $K$ to be a number field. For each place $v$ fix an embedding $\olK \subseteq \olK_v$. Then $\Gal(\olK_v/K_v) \subseteq \Gal(\olK/K)$. This gives us a short exact sequence resembling the one above:
\begin{align*}
	0 \to \prod_v\frac{E'(K_v)}{\phi E(K_v)} \to \prod_vH^1(K_v,E[\phi]) \to \prod_{ v }H^1(K_v,E)[\phi*] \to 0.
\end{align*}
These products just mean that we have an exact sequence
\begin{align*}
	0 \to \frac{E'(K_v)}{\phi E(K_v)} \to H^1(K_v,E[\phi]) \to H^1(K_v,E)[\phi*] \to 0
\end{align*}
for each place $v$. We also have the following commutative diagram with exact rows:
\begin{center}
\begin{tikzcd}
	0 \arrow[r]& \frac{E'(K)}{\phi E(K)}\arrow[d] \arrow[r,"\delta"]& H^1(K,E[\phi]) \arrow[rd] \arrow[d,"res_v"] \arrow[r] & H^1(K,E)[\phi*]\arrow[d,"res_v"] \arrow[r]& 0 \\
	
	0 \arrow[r]& \prod_v\frac{E'(K_v)}{\phi E(K_v)} \arrow[r]& \prod_vH^1(K_v,E[\phi]) \arrow[r] &\prod_{ v }H^1(K_v,E)[\phi*] \arrow[r] & 0. 
\end{tikzcd}	
\end{center}
This leads us to the definition of the \emph{Selma group}.
\begin{defn}\label{defn:selma-group}
	The $\phi$-Selma group is
	\begin{align*}
		S^{(\phi)}(E/K) &= \ker(\text{downward diagonal map above}) \\
						&= \ker\big(H^1(K,E[\phi])\to \prod_v H^1(K_v,E)\big) \\
						&= \{ \alpha \in H^1(K,E[\phi]) ~\mid~ \res_v(\alpha) \in \img(\delta_v) ~ \forall v\}.
	\end{align*}
	The \emph{Tate Shaferevich group is }
	\begin{align*}
		\text{look at picture and fill in}, weird disjoint union looking symbol with three vertical strokes.
	\end{align*}
\end{defn}
We get a short-exact sequence
\begin{align*}
	0 \to \frac{E'(K)}{\phi E(K)} \to S^{(\phi)}(E/K) \to \tshaf(E/K)[\phi_*] \to 0.
\end{align*}
Taking $\phi = [n]$ gives
\begin{align*}
	0 \to \frac{E(K)}{nE(K)}\to S^{(n)}(E/K) \to \tshaf(E/K)[n] \to 0.
\end{align*}
Rearranging the proof of weak Mordell-Weil gives
\begin{thm}\label{thm:S-is-finite}
	$S^{(n)}(E/K)$ is finite.
\end{thm}
\begin{prf}
	For $L/K$ a finite Galois extension there is an exact sequence
	\begin{align*}
		0 \to H^1(\Gal(L/K), E(L)[n]) \to{\inf} H^1(K,E[n]) \to{\res} H^1(L,E[n]).
	\end{align*}
	The first nonzero term above is finite, and $S^{(n)}(E/K) \to S^{(n)(E/L)}$ is induced by $\res$ since $S^{(n)}(E/K) \subseteq H^1(K,E[n])$ and $S^{(n)(E/L) \subseteq H^1(L,E[n])}$. Therefore, by extending our field, we may assume $E[n] \subseteq E(K)$ and hence $\mu_n \subseteq K$. This implies that $E[n] \cong \mu_n \times \mu_n$ as a $\Gal(\olK/K)$-module.

	Therefore $H^1(K,E[n]) \cong H^1(K,\mu_n) \times H^1(K,\mu_n)\cong K^*/(K^*)^n \times K^*/(K^*)^n$. Let
	\begin{align*}
		S = \text{primes of bad reduction for $E/K$} \cup \{v \mid n \infty\}.
	\end{align*}
	N.B. This is a finite set of places.
\end{prf}

\begin{defn}\label{defn:unramifiedoutsideS}
	The subgroup of $H^1(K,A)$ unramified outside $S$ is
	\begin{align*}
		H^1(K,A;S) = \ker\big(H^1(K,A)\ to \prod_{v\not\in S} H^1(K_v^{nr},A)\big)
	\end{align*}
	There is a commutative diagram with exact rows 
	\begin{center}
		<put commutative diagram here>
	\end{center}
	This map is surjective (the $x_n$ map) for all $v \not\in S$ (see Theorem 9.7 from class) therefore $\img(\delta_v) \subseteq \ker(\text{green downward map})$.
\end{defn}
\begin{lem}
	Let $\ker\big(H^1(K,\mu_n) \to H^1(K^{nr},\mu_n)\big) \subseteq \{x \in K^*/(K^*)^n ~ \mid ~ v(x) \equiv 0 ~ (\mod n)\}$.
	Therefore
	\begin{align*}
		S^{(n)}(E/K) &= \left\{\alpha \in H^1(K, E[n]) ~\mid ~ \res_v(\alpha) \in \img(\delta_v) ~ \forall v\right\} \\
					 &\subseteq H^1(K,E[n]; S) \\
					 &\cong H^1(K,\mu; S) \times H^1(K,\mu_n ; S) \\
					 &\cong K(S,n)\times K(S,n).
	\end{align*}
	But $K(S,n)$ is finite by Lemma 11.4, therefore $S^{(n)}(E/K)$ is finite.
\end{lem}

\begin{rmk}
	$S^{(n)A}(E/K)$ is finite and effectively computable. If is conjectured that $|\tshaf(E/K) < \infty$. This would imply that $\rank E(K)$ is effectively computable.
\end{rmk}

\section{Descent by cyclic isogeny}
Let $E$ and $E'$ be elliptic curves over a number field $K$, and let $\phi:E\to E'$ be an isogeny of degree $n$. Suppose $E'[\hat{\phi}] \cong \bZ/n\bZ$ as a Galois module $S\mapsto e_\phi(S,T)$. Short-exact sequence of $\Gal(\olK/K)$-modules
\begin{align*}
	0\to \mu_n \to E\to{\phi} E' \to 0.
\end{align*}
Long exact sequence
\begin{center}
	\begin{tikzcd}
		...\arrow[r]& E(K) \arrow[r,"\phi"] & E'(K)\arrow[r,"\delta"] \arrow[rd,"\alpha"] & H^1(K,\mu_n) \arrow[r] \arrow[d, "\cong"]& ... \\
					 & & & K^*/(K^*)^n & 
	\end{tikzcd}
\end{center}
\begin{thm}
	Let $f \in K(E')$ and $g \in K(E)$ with $\div(f) = n(T) - n(P)$ and $\phi^*f = g^n$. Then $\alpha(P) = f(P) ~ \mod(K^*)^n$ for all $P \in E'(K) \setminus \{0,T\}$.
\end{thm}
\begin{prf}
	Let $Q \in \phi^{-1}P$. Then $\delta(P)$ is represented by the cocycle $\sigma \mapsto \sigma Q - Q \in E[\phi] \cong \mu_n$.
	\begin{align*}
		e_\phi(\sigma Q - Q, T) &= \frac{g(rQ - Q + X)}{gX)} &\text{ for any  } x\in E\setminus \text{zeros and poles} \\
							   &=  \frac{g(\sigma Q)}{g(Q)} & x = Q \\
							   &= \frac{\sigma \sqrt[n]{f(P)}}{\sqrt[n]{f(P)}} & \text{N.B.} f(P) = g(Q)^n
	\end{align*}
	Therefore $\delta(P)$ is represented by the cocycle $\sigma \mapsto \frac{\sigma(\sqrt[n]{f(P)})}{\sqrt[n]{f(P)}}$. But $H^1(K,\mu_n) \cong K^*/(K^*)^n$, $\\big(\sigma \mapsto \frac{\sigma(\sqrt[n]{x})}{ \sqrt[n]{x} }\big) \leftmapsto x$. Therefore $\alpha(P) = f(P) ~\mod(K^*)^n$.
\end{prf}

\lecture{2022-03-14}{2022-03-14}
\begin{thm}\label{thm:16-1}
	Let $f \in K(E')$ and $g \in K(E)$ with $\div(f) = n(T) - n(0)$ and $\phi^*f = g^n$. Then there exists a group homomorphism $\alpha: E'(K) \to K^*/(K^*)^n$ with $\ker \alpha = \phi(E(K))$ and $\alpha(P) = f(P) \mod (K^*)^n$ for all $P \in E'(K)\setminus \{0,T\}$.
\end{thm}

\subsection{Descent by 2-isogeny}
$E:~ y^2 = x(x^2 ax + b)$

$E':~ y^2 = x(x^2+a'x + b')$ where $b(a^2 - 4ab) \neq 0$, $a' = -2a$ $b' = a^2 - 4b$. Let $\phi:E\to E',$ $(x,y) \mapsto \left(\left(\frac{x}{y}\right)^2, \frac{y(x^2 - b)}{x^2}\right)$. Then
\begin{align*}
	\hat{\phi} E'\to E, ~ (x,y) \mapsto \left(\frac{1}{4}\left(\frac{y}{x}\right), \frac{y(x^2 - b')}{8x^2}\right)^2.
\end{align*}
Then $E[\phi] = \{0,T\}$, $T = (0,0) \in E(K)$ and $E'[\hat{\phi}] = \{0,T'\}$, $T' = (0,0) \in E'(K)$.
\begin{prop}
	There is a group homomorphism
	\begin{align*}
		E'(K) \to K^*/(K^*)^2, ~ (x,y) \mapsto
		\begin{cases}
			x(K^*)^2 & \text{if } x \neq 0 \\
			b'(K^*)^2 & \text{if } x = 0
		\end{cases}
	\end{align*}
	with kernel $\phi E(K)$.
\end{prop}
\begin{prf}
	\textbf{Either} Apply Theorem (\ref{thm:16-1}) with $f = x \in K(E')$ and $g = \frac{y}{x} \in K(E)$ \textbf{or} do direct calculation, see example sheet 4. 
\end{prf}
Two maps
	\begin{align*}
		\alpha_E:& \frac{E(K)}{\hat{\phi}E'(K)} \hookrightarrow K^*/(K^*)^2 \\
		\alpha_{E'}:& \frac{E'(K)}{\phi E(K)} \hookrightarrow K^*/(K^*)^2.
	\end{align*}
\begin{lem}\label{lem:16-3}
	\begin{align*}
		2^{\rank E(K)} = \frac{|\img(\alpha_E)| \cdot |\img \alpha_{E'}}{4}.
\end{align*}
\end{lem}
\begin{prf}
	If
	\begin{align*}
		A \to{f} B \to{g} C
	\end{align*}
	is a homomorphism of abelian groups then there is an exact sequence
	\begin{align*}
		0 \to \ker(f) \to \ker(gf) \to{f} \ker(g) \to \coker(f) \to{g} \coker(gf) \to \coker(g) \to 0.
	\end{align*}
	Since $\hat{\phi}\phi = [2]_E$ we get an exact sequence
	\begin{align*}
		0 \to E(K)[\phi] \to E(K)[2] \to{\phi} E'(K)[\hat{\phi}] \to \frac{E'(K)}{\phi E(K)} \to{\hat{\phi}} \frac{E(K)}{2E(K)} \to \frac{E(K)}{\hat{\phi}E'(K)} \to 0.
	\end{align*}
	The leftmost nontrivial term above is isomorphic to $\bZ/2\bZ$, the third nontrivial term is also $\bZ/2\bZ$, the fourth is isomorphic to $\img \alpha_{E'}$, and the rightmost nontrivial term is $\img \alpha_E$.

	Therefore
	\begin{align*}\label{eq:1-lec23}
		\frac{|E(K)/2E(K)}{|E(K)[2]} = \frac{|\img \alpha_E| \cdot |\img \alpha_{E'}|}{2\cdot 2}.
	\end{align*}
	Mordell-Weil implies $E(K) \cong \Delta \times \bZ^r$ where $\Delta$ is a finite group, $r = \rank E(K)$.
	\begin{align*}
		\frac{E(K)}{2E(K)} \cong \frac{\Delta}{2\Delta} \times \left(\bZ/2\bZ\right)^r
	\end{align*}
	and $E(K)[2] \cong \Delta[2]$. Therefore $\frac{|E(K)/2E(K)|}{|E(K)[2]} = 2^r$. Taken with equation (\ref{eq:1-lec23}), this proves the result.
\end{prf}
\begin{lem}\label{lem:16-4}
	If $K$ is a number field and $a,b \in \cO_K$ then $\img(\alpha_E) \subseteq K(S,2)$ where $S = \{\text{primes dividing $b$}\}$.
\end{lem}
\begin{prf}
	Must show that if $x,y \in K,$ $y^2 = x(x^2 + ax + b)$ and $v_\frakp(b)$, then $v_\frakp(x) = 0 ~(\mod 2)$.

	Case $v_\frakp(x) < 0$, then Lemma 9.1 $\implies$ $v_\frakp(x) = -2r$ and $v_\frakp(y) = -3r$ for some $r \geq 1$.

	Case $v_\frakp(x) < 0$, then $v_\frakp(x^2 + ax + b) = 0 \implies $ $v_\frakp(x) = v_\frakp(y^2) = 2v_{\frakp}(y)$.
\end{prf}

\begin{lem}\label{lem:16-5}
	If $b_1b_2 = b$ then $b_1(K^*)^2 \in \img(\alpha_E)$ or equivalently $\omega^2 = b_1u^4 + au^2v^2 + b_2v^4$ is soluble for $u,v,w \in K$ not all zero.
\end{lem}
\begin{prf}
	If $b_1 \in (K^*)$ or $b_2 \in (K^*)^2$ then both conditions are satisfied. So we may assume $b_1,b_2 \not\in (K^*)^2$. Have $b_1(K^*) \in \img(\alpha_E) \iff$ there exists some $(x,y) \in E(K)$ such that $x = b_1t^2$ for some $t \in K^*$. This implies $y^2 = b_1t^2\left((b_1)^2 + ab_1t^2 + b\right)\implies \left(\frac{y}{b_1t}\right)^2 = b_1t^4 + at^2 + b/b_1$. So the $\omega^2$ equation above has a solution $u = t, v = 1, \omega = \frac{y}{b_1t}$.

	Conversely (simply perform same calculation in reverse), if $(u,v,\omega)$ is a solution to the $\omega$ equation above, then $uv \neq 0$ and $\left(b_1\left(\frac{u}{v}\right)^2,b_1\frac{u\omega}{v^3}\right) \in E(K)$.
\end{prf}

\begin{example}
	Take $K = \bQ$ and $E:~ t^2 = x^3 - x$, $a = 0$ and $b = -1$. Then $\img(\alpha_E) = \langle -1 \rangle \subseteq \bQ^*/(\bQ^*)^2$, $E':~ y^2 = x^3 + 4x$. $\img(\alpha_E') \subseteq \langle -1,2 \rangle \subseteq \bQ^*/(\bQ^*)^2$.

	Have 
	\begin{align*}
		b_1 &= -1 & \omega^2 = -y^4 - 4v^4 \\
		b_1 &= 2 & \omega^2 = 2u^4 + 2v^4 \\
		b_1 &= -2 & \omega -2u^4 - 2v^4.
	\end{align*}
	The first and third equations are insoluble over $\bR$, while the second has solution $(u,v,\omega) = (1,1,2)$. Therefore $\img(\alpha_{E'}) = \langle 2 \rangle \subseteq \bQ^*/(\bQ^*)^2$ and $2^{\rank E(\bQ)} = \frac{2 \cdot 2}{4} \implies \rank E(\bQ) = 0 \implies 1$ is not a congruent number.
\end{example}

\begin{example}
	$E:~ y^2 = x^3 + px$ with $p$ prime $p \equiv 5 ~(\mod 8)$. Let $b_1 = -1,$ $\omega^2 = -u^4 - pv^4$ insoluble over $\bR$. Therefore $\img(\alpha_E) = \langle p \rangle\subseteq \bQ^*/(\bQ^*)^2$.
\end{example}

\lecture{2022-03-16}{2022-03-16}
Last time had an elliptic curve $E:~ y^2=x(x^2 + ax + b)$, $\phi:E \to E'$ a 2-isogeny. Set $b_2 = b/b_1$ and
\begin{align*}\tag{\ast}
	w^2 = b_1u^4 + au^2v^2 + b_2 v^4
\end{align*} 
\begin{align*}\tag{\ast'}
	w^2 = b_1u^4 + a'u^2v^2 + b_2 v^4
\end{align*}
with $b_2 = b'/b_1$. Here, we additionally had that $a' = -2a$ and $b' = a^2 - 4b$. We then get an exact sequence
\begin{center}
	\begin{tikzcd}
		0 \arrow[r]& \frac{E'(\bQ)}{\phi E(\bQ)} \arrow[r] \arrow[rd,"\alpha_{E'}"]& S^{(\phi)}(E/\bQ) \arrow[r] \arrow[d,"\text{inc}"]& \tshaf(E/\bQ)[\phi^*] \arrow[r]& 0 \\
				   & & \bQ^*/(\bQ^*)^2 & &
	\end{tikzcd}
\end{center}
Then $\img(\alpha_E) = \{b_1(\bQ^*)^2 ~ \mid ~ (\ast) \text{ is soluble over } \bQ\}$ is a subset of $S^{(\phi)}(E/\bQ) = \{b_1(\Q^*)^2 ~\mid~ (\ast') \text{ is soluble over }\bR \text{ and over  } \bQ_p \text{ for all primes } p\}$. 

Fact: (Uses Ex Sheet 3, Question 9 and Hensel's lemma) If $a, b_1, b_2 \in \bZ$ and $p\not | 2b(a^2 - 4b)$ then $(\ast)$ is soluble over $\bQ$.

\begin{example}[Continued from last lecture]
	$y^2 = x^3 + px$ with $p$ prime and $p \equiv 5~ (\mod 8)$, then 
	\begin{enumerate}[(1)]
		\item $w^2 = 2u^4 - 2pv^4$
		\item $w^2 = -2u^4 + 2pv^4$
		\item $w^2 = pu^4 - 4v^4$.
	\end{enumerate}
	(1) and (2) are insoluble over $\bQ_p$ since $\left(\frac{2}{p}\right) = \left(\frac{-2}{p}\right) = -1$, e.g. if (2) had a solution with $u,v,w \in \bQ_p$ (not all zero) then without loss of generality $u,v \in \bZ_p$ are coprime.

	If $p | u$ then $p | w$ and then $p | v$, contradiction. Therefore
	\begin{align*}
		\rank E(\bQ) =
		\begin{cases}
			0 & \text{if (3) is insoluble over } \bQ \\
			1 & \text{if (3) is soluble over} \bQ
		\end{cases}.
	\end{align*}
	(3) is soluble over $\bQ_p$ since $\left(\frac{-1}{p}\right) = + 1$, so by Hensel's lemma, $-1 \in (\bZ_p^*)^2$. (3) is insoluble over $\bQ_2$ since $p - 4 \equiv 4 ~ (\mod 8)$ so by Hensel's lemma $p - 4 \in (\bZ_2^*)^2$. (3) is soluble over $\bR$ since $\sqrt{p} \in \bR$.
\end{example}

It is an \textbf{open conjecture} that $\rank E(\bQ) = 1$ for all primes $p \equiv 5 ~ (\mod 8)$.

\begin{example}[Lind]
	$E: ~y^2 = x^3 + 17x$. $\img(\alpha_E) = \langle 17 \rangle\subseteq \bQ^*/(\bQ^*)2$. Set $E':~ y^2 = x^3 - 68x$, then $b_1 = 2$ and $\omega^2 = 2u^4 - 34v^4$. Replace $w$ by $2w$ and divide by 2 to get
	\begin{align*}\label{eq:lec24-1}
		C: 2w^2 = u^4 - 17v^4.
	\end{align*}
\end{example}

\textbf{Notation:} 

\begin{align*}
	C(K) = \left\{(u,v,w) \in K^3 \setminus \{0\} \midd \text{satisfying equation (\ref{eq:lec24-1})}\right\}
\end{align*}
where $(u,v,w) \sim (\lambda u, \lambda v, \lambda^2 w)$ for all $\lambda \in K^*$.
\begin{itemize}
	\item $C(\bQ_2) \neq \emptyset$ since $\17 \in (\bQ_2^*)^2$
	\item $C(\bQ_{17}) \neq \emptyset$ since $2 \in (\bQ_{17}^*)^2$
	\item $C(\bR) \neq \emptyset$ since $\sqrt{2} \in \bR$.
\end{itemize}
Therefore $C(\bQ_p) \neq \emptyset$ for all places $v$ of $\bQ$. Suppose $(u,v,w) \in C(\bQ)$ with (wlog) u,v,w \in \bZ, $\gcd(u,v) = 1$, $w > 0$. If $17 | w$, then $17 | u$ and then $17 | v$. Contradiction since $u$ and $v$ assumed to by coprime. So if $p | w$ with $p$ and odd prime, then $p\neq 17$ and $\left(\frac{17}{p}\right) = 1$ which implies $\left(\frac{p}{17}\right) = \left(\frac{17}{p}\right) = 1$ by quadratic reciprocity. Also note: $\left(\frac{2}{17}\right) = 1$, therefore $\left(\frac{w}{17} = 1\right)$. 

But $2w^2 \equiv u^4 ~(\mod 17)$, hence $2 \in (\bF_{17}^*)^4 = \{\pm_1, \pm 4\}$. A contradiction. Therefore $C(\bQ) = \emptyset$, i.e. $C$ is a counterexample to the Hasse principle. It represents a nontrivial element of $\tshaf(E/\bQ)$.

\subsectiot{Birch Swinterton-Dyer Conjecture}
Let $E/\bQ$ be an elliptic curve.
\begin{defn}\label{defn:lec24-1}
	\begin{align*}
		L(E,s) = \prod_p L_p(E,s)
	\end{align*}
	where
	\begin{align*}
		L_p(E,s) = 
		\begin{cases}
			\left(1 - a_p p^{-s} + p^{1 - 2s}\right)^{-1} & \text{if $E$ has good reduction at $p$} \\
			(1 \pm p^{-s})^{-1} & \text{if $E$ has multiplicative reduction at $p$} \\
			1 & \text{if E has additive reduction at $p$}.
		\end{cases}
	\end{align*}
	Here $\# \tilde{E}(\bF_p) = p+1 - a_p$.
\end{defn}
Hasse's theorem implies $|a_p| \leq 2\sqrt{p}$ and so $L(E,s)$ converges for $\Re(s) > 3/2.$

\begin{thm}[Wiles, Breil, Conrad, Diamond, Taylor]\label{thm:modularity-theorem}
	$L(E,s)$ is the $L$-function of a weight 2 modular form and hence has an analytic continuation to all of $\bC$ (and a function equation $L(E,s) \leftrightarrow L(E,2 - s)$).
\end{thm}
Wiles proved the special case of the modularity theorem for semi-simple (semi-stable?) elliptic curves, which was good enough for Fermat's last theorem.

The weak BSD:
\begin{conj}[Weak BSD]\label{conj:weak-BSD}
	$\ord_{s=1}L(E,s) = \rank E(\bQ)A.$
\end{conj}
\begin{conj}[Strong BSD]\label{conj:strong-BSD}
	This says 
	\begin{align*}
		\lim_{s\to 1} \frac{1}{(s - 1)^r} L(E,s) = \frac{\Omega_E \cdot \Reg E(\bQ) \cdot |\tshaf(E/\bQ)|\cdot \prod_p c_p}{|E(\bQ)_{tors}|^2}
	\end{align*}
	where 
	\begin{itemize}
		\item $c_p$ is the Tamagawa number of $E/\bQ_p$, i.e. $c_p = [E)\bQ_p) : E_0(\bQ)_p]$
		\item $E(\bQ)/E(\bQ)_{tors} = \langle p_1,...,p_r \rangle$,
		\item $\Reg E(\bQ) = \det([P_i, P_j])_{i,j = 1,...,r}$, 
		\item $\Omega_E = \int_{E(\bR)} \frac{dx}{2y + a_1x + a_3}$,
		\item $[P,Q] = \hat{h}(P+Q) - \hat{P} - \hat{Q}$ 
		\item $a_i = $ coefficient of a global minimal Weiestrauss equation for $E/Q$.
	\end{itemize}
\end{conj}
\end{document}
