\input{../preamble.tex}

\begin{document}
\begin{center}
	\Large
	\begin{LARGE}
		D-Modules, Unit \textit{F}-Crystals, and Hodge Theory \\
	\end{LARGE}
	Definitions, Theorems, Remarks, and Notable Examples \\
	Isaac Martin \\
    Last compiled \today
\end{center}
\normalsize
\vspace{-2mm}
\hru

\tableofcontents
\newpage
\section{Questions}
\begin{question}\label{q:D-mod-why-care}
	What is the point of D-modules? Why do people care about them? What sort of questions do they answer? What insights do they provide?
\end{question}
\begin{question}\label{q:sheaf-of-k-linear-endo}
	Given a scheme $X$ over a field $k$, we define the sheaf $\sEnd_k(\cO_X) = \sHom_k(\cO_X,\cO_X)$ to be the sheaf of $k$-linear endomorphisms on $\cO_X$ as in \cite[Definition 0.3]{Gieseker75}. What I think this means is this: we have a structure morphism $X \to \Spec k$ and we can therefore think of $\cO_X$ as a $\cO_{\Spec k}$-module via the map $f^\sharp:\cO_{\Spec k}\to f_*\cO_X$. However, this is rather stupid, since this means $\End_k(\cO_X)$ is a $\cO_{\Spec k}$-module, which should mean it takes open sets from $\Spec k$. In the book \cite[Example A.4.2.]{D-mod_PS_RT} however, this sheaf takes open sets from $X$. What's going on here?
\end{question}

\begin{question}\label{q:why-Weyl-subalgebra-natural}
	How can one "naturally" make $A_m$ a subalgebra of $A_n$ when $m \leq n$? It seems like there are $n$ many subalgebras of $A_n$ isomorphic to $A_{n-1}$, for example.
\end{question}
\section{Some Non-Commutative Algebra}
$\cD$-modules requires non-commutative algebra. Necessary facts are found here.
\subsection{Filtered rings and modules}
This subsection follows Ginzburg's notes quite closely, see [BIBTEX SETUP, GINZBURG D-MODULES Page 3]. 

\begin{defn}[\emph{Filtered Ring}]\label{defn:filtered-ring}
	Let $A$ be an associative ring with unit. We call $A$ a \emph{filtered ring} if an increasing filtration $... \subset A_i \subset A_{i+1} \subset ...$ by additive subgroups is given such that
	\begin{enumerate}[(i)]
		\item $A_iA_j \subset A_{ij}$
		\item $1 \in A_{0},$
		\item $\bigcup A_i = A$, i.e. the filtration is \emph{exhausting}.
	\end{enumerate}
	Typically, either (a) $~\bN$ or (b) $~\bZ$ is chosen for the index set. In the former case $A$ is said to be \emph{positively filtered}. Note that (a) can be viewed as a special case of (b) by setting $A_{-1} = 0$. In the latter case we will consider the topology induced by the filtration by taking $\left\{A_i\right\}_{i\in \bZ}$ to be the base of open sets, and we then impose two additional conditions:
	\begin{enumerate}
		\item[(iv)] $\bigcap A_i = \{0\}$, i.e. the topology defined by $\{A_i\}$ is \emph{separating}
		\item $A$ is complete with respect to this topology.
	\end{enumerate}

	\noindent Finally, we denote by $\gr A$ the associated graded ring $\bigoplus A_i / A_{i-1}$. 
\end{defn}

\newpage
\section{Differential Operators and D-Modules}
\begin{defn}[Quasi-coherent \#1]\label{defn:quasi-coh-Bernstein}
	Fix $X$ a scheme over $k$, $\cO_X$ the structure sheaf, $\cF$ a sheaf of $\cO_X$-modules. We call $\cF$ a \textit{quasi-coherent} sheaf of $\cO_X$-modules (or simply an $\cO_X$-modules) if it satisfies the condition
	\begin{center}
		If $U\subseteq X$ an open affine, $f \in \cO_X(U),$ and $U_f = \left\{u \in U \midd f(u) \neq 0\right\}$,
	\end{center}
	then $\cF(U_f) = \cF(U)_f = \cO_X(U_f)\otimes_{\cO_X(U)} \cF$.
\end{defn}

\begin{defn}[Quasi-coherent \#2]\label{defn:quasi-coh-Hart}
	Let $(X,\cO_X)$ be a scheme. A sheaf of $\cO_X$-modules $\cF$ is quasi-coherent if $X$ can be covered by affine opens $U_i = \Spec A_i$ such that for each $i$ there exists an $A_i$ module $M_i$ with $\cF|_{U_i} \cong \tilde{M_i}$. We say $\cF_i$ is coherent if each $M_i$ can be taken to be finitely generated.
\end{defn}
\begin{rmk}\label{rmk:sheaf-ass-to-M}
	If $A$ is a ring and $M$ an $A$-module, the sheaf associated to $M$ is denoted by $\tilde{M}$ and is formed as follows. For each $\frakp \in \Spec A$, $M_\frakp = A_{\frakp} \otimes_A M$ is the localization with respect to $\frakp$. Given an open set $U \subseteq \Spec A$, define
	\begin{align*}
        \tilde{M}(U) = \left\{s:U\to \bigsqcup_{\frakp \in U} M_\frakp \midd s(\frakp) \in M_\frakp, \text{ and locally } s = \frac{m}{f}, m \in M, f \in A\right\}.
	\end{align*}
	More verbosely, this last condition means that for each $\frakp \in U$ there is a neighborhood $V \subseteq U$ of $\frakp$ such that for each $\frakq \in V,$ $f\not\in \frakq$ and $s(\frakq) = \frac{m}{f} \in M_\frakq$.

	Alternatively, one may define
	\begin{align*}
	    \tilde{M}(U_f) = M_f,
	\end{align*}
	and then
	\begin{align*}
	    \tilde{M}(U) = \varinjlim_{U_f \subseteq U} \tilde{M}(U_f).
	\end{align*}
	Note that $U_f$ is implied to be a distinguished open in one of the $U_i$, so really we need to take the limit above over all $U_f$ in all $U_i$ which intersect $U$ nontrivially. This is a non-issue if $U$ is affine.
\end{rmk}
\begin{lem}\label{lem:ginzburg-characterization-of-quasi-coh}
	The following are equivalent conditions for $\cF$ a sheaf of $\cO_X$ modules:
	\begin{enumerate}[(a)]
		\item $\cF$ is the direct limit of its coherent subschemes
		\item For any Zariski open affine subset $U\subseteq X$ and any $f \in \cO(U)$ one has $\Gamma(U_f, \cF) = \Gamma(U,\cF)_f$.
	\end{enumerate}
	A \textit{quasi-coherent} sheaf is then one which satisfies these conditions. 
\end{lem}
\begin{lem}[Noether Normalization Lemma]\label{lem:Noeth-Normal}
	Let $k$ be a field, $A$ a finitely generated $k$-algebra. Then there exists algebraically independent elements $y_1, ..., y_d$ in $A$ for some positive $d$ such that $A$ is finitely generated as a module over $k[y_1,...,y_n]$.
\end{lem}

\begin{rmk}\label{rmk:Noeth-Norm-gives-coord-def-of-diff-op}
	The Noether normalization lemma provides a way to define differential operators using a manifold-esque coordinate approach. I prefer the following coordinate-free approach provided by Gr\"othendieck, however.
\end{rmk}
\begin{defn}[Differential Operators]\label{defn:diff-ops}
	Let $A$ be a commutative ring. For any pair of $A$-modules $M$, $N$ we define the module $\Diff_A^{k}(M,N)$ inductively as follows:
	\begin{enumerate}[(i)]
		\item $\Diff_A^{0}(M,N) = \Hom_A(M,N)$
		\item $\Diff_A^{k+1}(M,N) = \left\{\text{ additive maps } u:M\to N \midd \forall a \in A, (a u - u a) \in \Diff^{k}_A(M,N)\right\}$
	\end{enumerate}
	It follows from the definition that $\Diff_{A}^{k}(M,N) \subset \Diff_A^{k+1}(M,N)$. We define
	\begin{align*}
		\Diff_A(M,N) := \bigcup_{k} \Diff_A^k(M,N).
	\end{align*}
	In the case that $M + N$, we write $\Diff_A(M)$ and note that it is a filtered almost commutative ring.
\end{defn}
The case in which we will be most interested is when $M = N = A$, i.e. when we consider $A$ to be a ring over itself. Let's repeat the above construction for this case.

\begin{defn}\label{def:order-of-operator}
	Let $A$ be a commutative $K$-algebra for $K$ a characteristic 0 field. Let $D \in \End(R)$. We define the \textbf{order} of $D$ inductively.
	\begin{itemize}
		\item $\ord(D) = 0$ if $[a,D] = -[D,a] = 0$ for all $a \in A$.
		\item $\ord(D) = n \in \bZ_{\geq 0}$ if $\ord(D) \neq k$ for all $k < n$ and if $\ord([a,D]) = k_a$ for some $k_a < n$ for each $a \in A$. 
	\end{itemize}
	The set $D^n(R)$ is defined to be the  $K$-vector space of of all operators of order $ \leq n$. 
\end{defn}
\begin{defn}\label{def:derivation}
	A \textbf{derivation} $D \in \End(R)$ is an operator which satisfies the Leibniz rule
	\begin{align*}
		D(ab) = aD(b) + D(a)b
	\end{align*}
	for every $a,b \in A$. The set of all derivations $\Der_K(A) \subseteq \End_K(A)$ is a $K$-vector space and a left $A$-module under the action $(a\cdot D)(b) = a(D(B))$.
\end{defn}
All derivations are order 1 operators. As one might hope, they're actually \emph{all} order 1 operators.
\begin{lem}\label{lem:derivations-are-order-one}
	$D^1(A) = \Der_K(R) + R$. (See proof in \emph{A Primer on $D$-modules} page 21.)
\end{lem}
We can now define the ring of differential operators on $A$.
\begin{defn}\label{defn:ring-of-diff-ops}
	Let $A$ be a $K$-algebra with $K$ a characteristic zero field. The set of all finite order operators on $A$ forms a noncommutative ring with pointwise addition and composition as multiplication. We denote this ring by $D(A)$ and we call it the \textbf{ring of differential operators:}
	\begin{align*}
		D(A) = \bigcup_{n \in \bZ_{\geq 0}} D^n(A) \subseteq \End_K(A).
	\end{align*}
\end{defn}
It's obvious that the addition of two finite order operators yields a differential operator with order equal to the maximum order of the two summands, it's \emph{not} obvious that the composition of two finite order operators yields a finite order operator. We therefore require the following proposition for this definition to work:
\begin{prop}\label{prop:composition-of-operators}
	If $D \in D^n(A)$ and $D' \in D^m(A)$ then $D\circ D', D' \circ D \in D^{n+m}(A)$.
\end{prop}



\subsection{First examples of Differential Operators}
\begin{example}\label{exp:Weyl-algebra}
	Let $K$ be a field of characteristic zero and recall that $K[x_1,...,x_n]$ is an infinite dimensional $K$ vector space. We define $\hat{x_i}, \partial_i \in \End_K(K[x_1,...,x_n])$ by $\hat{x_i} f \mapsto x_i\cdot f$ and $\partial_i f \mapsto \frac{\partial f}{\partial x_i}$. One can then check that $[\partial_j, \hat{x_j}] = \partial_i \hat{x_j} - \hat{x_j}\partial_i = \delta_{ij}\id$ where $\id$ is the identity operator on $K[x_1,...,x_n]$ and $\delta_{ij}$ is the Kronecker delta. In other words,
	\begin{align*}
		[\partial_i,\hat{x_i}](f) = f \buffword{1em}{and} [\partial_i,\hat{x_j}](f) = 0
	\end{align*}
	when $i \neq j$. This is quite easy to check for an arbitrary polynomial but is nonetheless quite magical:
	\begin{align*}
		\partial_x \left(x\cdot (3x^2 + 2y)\right) &= 9x^2 + 2y, \\
		x \cdot \left(\partial_x(3x^2 + 2y)\right) &= 6x^2, \\
		(\partial_x \cdot \hat{x} - \hat{x}\cdot \partial_x)(3x^2 + 2y) = 3x^2 + 2y,
	\end{align*}
	but
	\begin{align*}
		\partial_x \left(y\cdot(3x^2 + 2y)\right) - y\cdot \left(\partial_x(3x^2 + 2y)\right) = 6xy - 6xy = 0.
	\end{align*}
	\begin{defn}\label{def:Weyl-alg}
		The $n$th Weyl algebra of $K$ is the $2n$-dimensional $K$-subalgebra of $\End_K(K[x_1,...,x_n])$ generated by $\hat{x_1},...,\hat{x_n},\partial_1,...,\partial_n$, and is denoted by $A_n(K)$ or $A_n$ when the field is known. We let $A_0(K) = K$.

		Note also that for $m \leq n$, we can make $A_m$ a subalgebra of $A_n$ in a "natural way".
	\end{defn}
\end{example}

\subsection{D-Modules}
\begin{defn}\label{def:D-modules}
	A $\cD$-module is a sheaf over the sheaf $\cD_X$ of regular differential operators over a variety (scheme, manifold, analytic complex manifold) which is quasi-coherent as an $\cO_X$-module.
\end{defn}


\section{Berstein-Sato Polynomial}
\begin{thm}[Bj\"ork, Kashiwara]\label{thm:b-function-theorem}
	Let $X$ be a smooth variety over the complex numbers and let $f$ be a non-invertible regular function on $X$ (i.e. a locally rational function whose numerator is non-invertible). There exists a polynomial $b(s) \in \bC[s]$ and a polynomial $P(s) \in \cD_X[s]$ whose coefficients are differential operators on $X$, such that the relation
	\begin{align*}
		P(s)f^{s+1} = b(s)\cdot f^s
	\end{align*}
	holds formally in the $\cD_X$-module $\cO_X[\frac{1}{f},s] \cdot f^s$. Here, $f^{s+1}$ stands for $f\cdot f^s$.
\end{thm}


\newpage
\printbibliography

\end{document}
