\input{../preamble.tex}

\usepackage{tikz}
\usetikzlibrary{positioning,calc,intersections,through,backgrounds, shapes.geometric, decorations.markings,arrows}

\def\sset{\subseteq}
\def\iso{\cong}
\def\gend#1{\langle #1\rangle}

\newcommand{\rightoverleftarrow}{%
  \mathrel{\vcenter{\mathsurround0pt
    \ialign{##\crcr
      \noalign{\nointerlineskip}$\longrightarrow$\crcr
      \noalign{\nointerlineskip}$\longleftarrow$\crcr
    }%
  }}%
}

\newcommand\makesphere{} % just for safety
\def\makesphere(#1)(#2)[#3][#4]{%
  % Synopsis
  % \makesphere[draw options](center)(initial angle:final angle:radius)
  \shade[ball color = #3, opacity = #4] #1 circle (#2);
  \draw #1 circle (#2);
  \draw ($#1 - (#2, 0)$) arc (180:360:#2 and 3*#2/10);
  \draw[dashed] ($#1 + (#2, 0)$) arc (0:180:#2 and 3*#2/10);
}
% same thing as makesphere but places white background behind
\newcommand\altmakesphere{} % just for safety
\def\altmakesphere(#1)(#2)(#3)[#4][#5]{%
  % Synopsis
  % \makesphere[draw options](center)(initial angle:final angle:radius)
  \draw [fill=white!30] #1 circle (#2);
  \shade[ball color = #4, opacity = #5] #1 circle (#2);
  \draw #1 circle (#2);
  \draw ($#1 - (#2, 0)$) arc (180:360:#2 and 3*#2/10);
  \draw[dashed] ($#1 + (#2, 0)$) arc (0:180:#2 and 3*#2/10);
  \node at #1 {#3};
}

\begin{document}
\pagestyle{empty}
	\LARGE
\begin{center}
	Algebraic Topology Homework 2 \\
	\Large
	Isaac Martin \\
    Last compiled \today
\end{center}
\normalsize
\vspace{-2mm}
\hru

\tchap{Problems from 1.1}
\begin{homework}[e]
    \prob[\textsc{Exercise 2.}] Show that the change of basepoint homomorphism $\beta_h$ depends only on the homotopy class of $h$.
    \begin{prf}
        Let $X$ be a topological space with $x_0, x_1 \in X$ and suppose $h,g: [0,1] \to X$ are homotopic paths such that $h(0) = g(0) = x_1$ and $h(1) = g(1) = x_0$. We would like to show that $\beta_h = \beta_g$, i.e. that $h$ and $g$ both induce the same homomorphism $\pi_1(X,x_0) \to \pi_1(X,x_1)$. Change of basepoint homomorphisms are isomorphisms by Proposition 1.5 in Hatcher, so this is equivalent to showing that $\beta_h \circ \beta_g^{-1} = \beta_g^{-1}\circ \beta_h = \id$. But since $\beta_g^{-1} = \beta_{\olg}$, this is a simple calculation. For any $[f] \in \pi_{1}(X,x_0)$, we have
        \begin{align*}
            \beta_h\beta_{\overline{g}}([f]) = \beta_h([\olg \cdot f\cdot g]) = [h \cdot \olg \cdot f \cdot g \cdot \olh] = [f]
        \end{align*}
        since $h \simeq g$, and similarly
        \begin{align*}
            \beta_{\olg}\beta_{h}([f]) = \beta_{\olg}([h \cdot f\cdot \olh]) = [\olg \cdot h \cdot f \cdot \olh \cdot g] = [f].
        \end{align*}
        This means $\beta_{\olg}$ is the inverse of both $\beta_g$ and $\beta_h$, and by the uniqueness of inverses, we conclude $\beta_h = \beta_g$.
    \end{prf}
	\prob[\textsc{Exercise 5.}] Show that for a space $X$, the following three conditions are equivalent:
	\begin{itemize}
		\item[(a)] Every map $S^1 \rightarrow X$ is homotopic to a constant map, with image a point.
		\item[(b)] Every map $S^1 \rightarrow X$ extends to a map $D^2 \rightarrow X$.
		\item[(c)] $\pi_1 (X,x_0) = 0$ for all $x_0 \in X$.
	\end{itemize}
	Deduce that a space $X$ is simply-connected iff all maps $S^1 \rightarrow X$ are homotopic. [In this problem, 'homotopic' means 'homotopic without regard to basepoints.']
	\begin{prf} We prove the following chain of implications:

	\vspace{1em}
		  $(a) \implies (b):~$ Suppose that for every map $S^1 \rightarrow X$ is nullhomotopic to a function $c_{x_0} : S^1 \rightarrow X$ where $c_{x_0}(x) = x_0$, i.e. suppose there exists a homotopy $f_t:S^1 \rightarrow X$ where $f_1 = c_{x_0}$ and where $f_0$ is any map $S^1 \rightarrow X$. Since $(D^2, S^1)$ is a CW-pair, it has the homotopy extension property, and thus $f$ extends to $f': D^2 \times [0,1] \rightarrow X$ where $f'|_{S^1} = f$. Thus every map $S^1 \rightarrow X$ extends to $D^2 \rightarrow X$.
		  
		  $(b) \implies (c):~$ Suppose that $[f] \in \pi_1(X,x_0)$. $f(0) = f(1) = x_0$, meaning that we can interpret $f$ instead as a function $f:S^1 \rightarrow X$. This $f$ can be extended to a function $f': D^2 \rightarrow X$, but since $D^2$ is contractible, $f'$ is nullhomotopic. Thus, $f$ is also nullhomotopic, and $[f] = [0]$. This is true of any arbitrary $[f] \in \pi_1(X,x_0)$, and so we know that $\pi_1(X,x_0)$ is trivial. Noticing that the choice of $x_0$ was arbitrary, we conclude that $\pi_1(X) = 0$.
		  
		  $(c) \implies (a):~$This argument is very similar to the previous one. Each map $S^1 \rightarrow X$ can be interpreted as a loop in $X$, and since we assume $\pi_1(X) = 0$, every loop is homotopic to the trivial map. Thus every map $S^1 \rightarrow X$ is homotopic to a constant map in $X$.
		  
		  \bigskip
		  
		  A space $X$ is "simply connected" if and only if it is path connected and $\pi_1(X) = 0$. As we just showed, this is equivalent to "$X$ is path connected and every $S^1 \rightarrow X$ is nullhomotopic". If we interpret two maps $f,g:S^1 \rightarrow X$ as loops in $X$ with basepoints $x_0$ and $x_1$ respectively, then $f$ and $g$ are homotopic in $X$ by the homotopy that first shrinks $f$ to the constant map $c_{x_0}$, moves $x_0$ along a path to $x_1$, and finally deforms $c_{x_1}$ into $g$. Thus, any two maps $S^1 \rightarrow X$ are homotopic. We conclude that $X$ is simply connected if and only if all maps $S^1 \rightarrow X$ are homotopic.
    \end{prf}

\prob[\textsc{Exercise 6.}] We can regard $\pi_1 (X,x_0)$ as the set of basepoint-preserving homotopy classes of maps $(S^1,s_0) \rightarrow (X,x_0)$. Let $[S^1,X]$ be the set of homotopy classes of maps $S^1 \rightarrow X$ with no conditions on basepoints. Thus there is a natural map $\Phi: \pi_1(X,x_0) \rightarrow [S^1,X]$ obtained by ignoring basepoints. Show that $\Phi$ is onto if $X$ is path-connected, and that $\Phi([f]) = \Phi([g])$ iff $[f]$ and $[g]$ are conjugate in $\pi_1(X,x_0)$. Hence $\Phi$ induces a one-to-one correspondence between $[S^1,X]$ and the set of conjugacy classes in $\pi_1(X)$, when $X$ is path-connected. 

\begin{prf}
We first show that $\Phi$ is onto. Let $[f]$ be a member of $[S^1, X]$. Since every map $S^1 \rightarrow X$ can be regarded as a loop in $X$, $f$ is a loop based at some point $x_1 \in X$. Because $X$ is path connected, there must be some path $\gamma$ from $x_1$ to $x_0$. The map $g = \gamma \cdot f \cdot \overline{\gamma}$ is a continuous path that begins and ends at $x_0$, and is therefore a loop around $x_0$. Since $f$ and $g$ are homotopic, $[g] = [f]$. Since $[g] \in \pi_1(X,x_0)$, $\Phi([g]) = [f]$. We conclude that $\Phi$ is surjective. 

\bigskip

We now show that $\Phi([f]) = \Phi([g])$ if and only if $[f]$ and $[g]$ are conjugates. We show the forward implication first.

Assume that $\Phi([f]) = \Phi([g])$. This means that $f$ and $g$ are in the same equivalence class of $[S^1, X]$, so there must exist a homotopy $\varphi : S^1 \times [0,1] \rightarrow X$, where $\varphi_0 = f$ and $\varphi_1 = g$. The induced homeomorphisms $\varphi_{0*}$ and $\varphi_{1*}$ then satisfy $\varphi_{0*} = \beta_h \varphi_{1*}$, where $\beta_h : \pi_1(X, \varphi_1(s_0)) \rightarrow \pi_1(X, \varphi_0(s_0))$ and $h$ is the loop $\varphi_t(s_0)$. This means 

\begin{equation*}
    \varphi_{0*}([1]) = [g \cdot 1] = [g] = \beta_h\varphi_{1*}([1]) = \beta_h([f]) = [hf\overline{h}]
\end{equation*}
where $[1]$ is the equivalence class isomorphic to $1 \in \bZ$.

\bigskip

We now show the reverse implication. Assume that $[g] = [h][f][\overline{h}]$ in $\pi_1(X, x_0)$, i.e. assume $[f]$ and $[g]$ are conjugates. We want to show that $[f] = [g]$, or that $[f] = [h][f][\overline{h}]$. Consider the following function:
\begin{equation*}
    F: [0,1] \times S^1 \rightarrow X 
    \hspace{30pt}
    F_s(t) = 
    \begin{cases}
        h(3t + s) & 0 \leq t \leq \frac{1-s}{3} \\
        f\left(\frac{3}{1+2s}(t - \frac{1 - s}{3})\right) & \frac{1 - s}{3} \leq t \leq \frac{2+s}{3} \\
        \overline{h}\left(3(t - \frac{2 + s}{3})\right) & \frac{2 + s}{3} \leq t \leq 1
    \end{cases}
\end{equation*}

Since $F_0 = h \cdot f \cdot \overline{h}$, $F_1 = f$, and $F$ is continuous by the pasting lemma, $F$ is a homotopy between $f$ and $h \cdot f \cdot \overline{h}$. Thus, $[f] = [h][f][\overline{h}]$ in $[S^1, X]$ and we conclude that $\Phi([g]) = \Phi([f])$.

\end{prf}
\prob[\textsc{Exercise 10.}]From the isomorphism $\pi_1(X\times Y, (x_0,y_0)) \approx \pi_1(X,x_0) \times \pi_1(Y,y_0)$ it follows that loops in $X \times \{y_0\}$ and $Y \times \{x_0\}$ represent commuting elements of $\pi_1\left(X \times Y, (x_0,y_0)\right)$. Construct an explicit homotopy demonstrating this.

\begin{prf}
    Let $[f]$ and $[g]$ be loops based at $x_0$ in $X$ and $y_0$ in $Y$, respectively. Next consider the following homotopies:
    \begin{equation*}f_t(s) = 
        \begin{cases}
            x_0 & 0 \leq s \leq \frac{t}{2} \\
            f(2s) & \frac{t}{2} \leq s \leq \frac{1+t}{2} \\
            x_0 & \frac{1 + t}{2} \leq s \leq 1
        \end{cases}
    \end{equation*}
    and 
    \begin{equation*}g_t(s) = 
        \begin{cases}
            y_0 & 0 \leq s \leq \frac{t}{2} \\
            g(2s) & \frac{t}{2} \leq s \leq \frac{1+t}{2} \\
            y_0 & \frac{1 + t}{2} \leq s \leq 1
        \end{cases}
    \end{equation*}
Here $f_0$ is the path that transverses $f$ and then stays at $x_0$. $f_1$ is the path that stays at $x_0$ for half the interval and then transverses $f$. $g_0$ is the path that stays at $y_0$ and then transverses $g$, and finally $g_1$ is the path that transverses $g$ and then stays at $y_0$.

Next since $\pi_1(X,x_0) \times \pi_1(Y,y_0) \approx \pi(X \times Y, (x_0,y_0))$ and $h_t(s) = (f_t(s),g_t(s))$ is an element of $\pi_1(X,x_0) \times \pi_1(Y,y_0)$, then $f_0 \cdot g_0 \simeq f_1 \cdot g_1$. However, since $f_0 \cdot g_0 \simeq f \cdot g$ and $f_1 \cdot g_1 \approx g \cdot f$, we conclude that $f\cdot g \simeq g\cdot f$.
\end{prf}

\prob[\textsc{Exercise 15.}] Given a map $f:X\to Y$ and a path $h:I\to X$ from $x_0$ to $x_1$, show that $f_*\beta_h = \beta_{fh}f_*$.
\begin{prf}
    Let $[\alpha] \in \pi_1(X,x_1)$ be an arbitrary equivalence class of loops based at $x_1$. By definition,
    \begin{align*}
        f_*(\beta_h([\alpha])) = f_*[h\cdot \alpha \cdot \olh] = [f\circ (h\cdot \alpha \cdot \olh)]
    \end{align*}
    and 
    \begin{align*}
        \beta_{fh}(f_*([\alpha])) = \beta_{fh}([f\circ \beta]) = [(f\circ h)\cdot (f\circ \alpha)\cdot \ol{(f\circ h)}].
    \end{align*}
    However, up to a possible reparameterization, $(f\circ h)\cdot (f\circ \alpha)\cdot \ol{(f\circ h)}$ and $f\circ (h\cdot \alpha \cdot \olh)$ are identical paths on $Y$. In fact, if we perform concatenation from consistently, then they are identical without \emph{any} reparameterization.
    
    To see this, concatenate from right to left without loss of generality. For $t \in [0,1/2]$ we have
    \begin{align*}
      f\circ (h\cdot (\alpha\cdot \olh))(t) = f(h(t)) = (f\circ h) \cdot \big((f\circ \alpha) \cdot \ol{(f\circ h)}\big)(t),
    \end{align*}
    and we have something similar for $t \in [1/2,3/4]$ and $t \in [3/4,1]$. As the representatives of the resulting equivalence classes above are homotopic, the diagram shown by Hatcher commutes.
\end{prf}

\prob[\textsc{Exercise 16.}]$ $
\begin{enumerate}[(a)]
  \item $X = \bR^3$ with $A$ any subspace homeomorphic to $S^1$.
  \item $X = S^1 \times D^2$ with $A$ its boundary torus $S^1\times S^1$.
  \item $X = S^1\times D^2$ with $A$ the interlocked circle in the solid torus.

\end{enumerate}
\begin{prf}
  \begin{enumerate}[(a)]
    \item The space $\bR^3$ contracts to the origin via the homotopy $F:\bR^3 \times I \to \bR^3$ defined $F_t(x) = (1 - t)x$, and hence has trivial fundamental group. By theorem 1.7, the circle $S^1$ has fundamental group isomorphic to $\bZ$. Therefore there exists no inclusion $\pi_1(S^1,x_0) \to \pi_1(\bR^3,x_0)$, and hence by Proposition 1.17 in Hatcher, there exists no retraction of $X$ onto a circle.

    \item In this case, we have that $\pi_1(X,x_0) \cong \bZ$ by Proposition 1.12 and $\pi_1(A,x_0) \cong \bZ\times \bZ$ by the same proposition. Note that because both of these spaces are path connected, the choice of basepoint does not matter. As any map $\bZ$-linear morphism $\bZ^2 \to \bZ$ will have nonzero kernel, there does not exist an injection $\bZ^2 \to \bZ$, and hence there is no retraction $r:A\to X$.

    \item The subset $A\subset X$ is contractible by pulling the two ends of the loops through each other. Thus, the map $\iota_*:\pi_1(A,x_0) \to \pi_1(X,x_0)$ induced by the inclusion $\iota:A\to X$ takes every loop in $A$ to something homotopic to the trivial loop in $X$. This means that the induced map $\iota_*$ is trivial, and in particular is not injective, hence by Proposition 1.17 $X$ does not retract onto $A$.
    \item Suppose we did have a retraction $r:X\to A$. In that case, the composition
      \begin{align*}
        D^2 \hookrightarrow X \xrightarrow{r} A \to S^1
      \end{align*}
      would also be a retraction. However, this is impossible, as $D^2$ is contractible while $S^1$ is not. Hence there is no such retraction $r$.
    \item 
    \item Since $X$ deformation retracts onto its central circle (a fact we used on the last homework) both $X$ and $A$ have fundamental group isomorphic to $\bZ$. Let $x_0 \in A \subset X$ be a basepoint, and choose generators $\gamma \in \pi_1(A,x_0)$ and $\lambda \in \pi_1(X,x_0)$ for the fundamental groups. The image $\iota_*([\gamma])$ of $[\gamma]$ under the map induced by the inclusion $\iota:A\to X$ is then equal to $2[\lambda]$ since traversing around the boundary circle corresponds to traversing around the central circle twice. Hence the induced map $\iota_*$ is really a map $\bZ\to \bZ$ which sends $2\mapsto 1$. This is impossible, as $2 \mapsto 1$ is not possible for such a group homomorphism as all $\bZ$-linear maps $\bZ\to \bZ$ are defined $1 \mapsto nz$ for some $n \in \bZ$. Hence, there is no retraction of $X$ onto the boundary $A$.
  \end{enumerate}
\end{prf}

\newpage

\tchap{Problems from 1.2}
\prob[\textsc{Exercise 3.}] Show that the complement of a finite set of points in $\bR^n$ is simply-connected if $n \geq 3$.

\begin{prf}
Let $\{x_1,...,x_k\}$ be a finite collection of points in $\bR^n$. We can place a ${n-1}$ sphere around every point removed from $\bR^n$, ensuring that the radius is small enough to contain only one "hole", and connect every sphere with another via a straight path. Call this space $X$. Just as $\bR^n$ with a single point removed deformation retracts to $S^{n-1}$, I claim that $\bR^n - \{x_1,...,x_k\}$ deformation retracts onto $X$. Every point inside of one of the spheres retracts onto the sphere, and choosing to map every point in $\bR^n$ that is not in $X$ onto the nearest point in $X$ retains continuity. 

Every path-connected open set on this surface has the trivial fundamental group, so by Van Kampen's theorem, the fundamental group of $X$ is also trivial. Since $\bR^n - \{x_1,...,x_k\}$ deformation retracts onto $X$, we conclude that $\pi_1(\bR^n - \{x_1,...,x_k\}) = [1]$. 
\end{prf}
\bigskip

\begin{tikzpicture}
  \coordinate (A) at (15:0.5cm);
  \coordinate (B) at (3,1);
  \coordinate (S1) at (0,0);
  \coordinate (S2) at (3,1);
  \coordinate (S3) at (3,-0.5);
  \coordinate (S4) at (9,-0.3);
  \coordinate (S5) at (7,0.5);
  \coordinate (S6) at (12,0);
  \draw (S1) -- (S2) -- (S3) -- (S4) -- (S5);
  \draw (S4) -- (S6);
  \altmakesphere((S1))(0.7)(\textbullet $x_1$)[gray!10][0.4]
  \altmakesphere((S2))(0.45)(\textbullet $x_2$)[gray!10][0.4]
  \altmakesphere((S3))(0.8)(\textbullet $x_3$)[gray!10][0.4]
  \altmakesphere((S4))(0.75)(\textbullet $x_4$)[gray!10][0.4]
  \altmakesphere((S5))(0.6)(\textbullet $x_5$)[gray!10][0.4]
  \altmakesphere((S6))(0.6)(\textbullet $x_6$)[gray!10][0.4]
  \node [] at (6,-2) {\emph{The space $X$ defined above}};
\end{tikzpicture}

\end{homework}
\end{document}
